Geometric models reveal behavioral and neural signatures of transforming naturalistic experiences into episodic memories

### 2020-09-03 paper/main.tex
removed all uses of the word experiment in reference to our study and chen et al, as per reviewer 3 request

\cite{ChenEtal17} collected hand-annotated information about each of 1,000 (manually identified) scenes spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening, the location where the scene took place, the names of any characters on the screen, and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the vocabulary for the topic model.  We then concatenated the sets of words across all features contained in overlapping sliding windows of (up to) 50 scenes, and treated each window as a single document for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the episode (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  We note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06}, in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched (interpolated) the resulting windows-by-topics matrix to match the time series of the 1,976 fMRI volumes collected as participants viewed the episode.

@@

\begin{figure}[tp]
  \centering
  \vspace*{-1cm}
  \includegraphics[width=.8\textwidth]{figs/precision_distinctiveness_detail}
  \caption{\small \textbf{Precision reflects the completeness of recall, whereas distinctiveness reflects recall specificity.} \textbf{A.} Recall precision by episode event.  Grey violin plots display kernel density estimates for the distribution of recall precision scores for a single episode event.  Colored dots within each violin plot represent individual participants' recall precision for the given event.  episode events are ordered along the $x$-axis by the average precision with which they were remembered.  \textbf{B.} The set of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising an example episode event (22) identified by the HMM.  Each action or feature is highlighted in a different color.  \textbf{C.} Excerpts from the most precise (P17) and least precise (P6) participants' recalls of episode event 22.  Descriptions of specific actions or features reflecting those highlighted in Panel B are highlighted in the corresponding color.  The text highlighted in gray denotes a (rare) false recall.  The unhighlighted text denotes correctly recalled information about other episode events.  \textbf{D.} Recall distinctiveness by episode event.  Kernel density estimates for each episode event's distribution of recall distinctiveness scores, analogous to Panel A.  \textbf{E.} The sets of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising episode events described by the example participants in Panel F.  Each event's text is highlighted in a different color.  \textbf{F.} The sentences comprising the most distinctive (P9) and least distinctive (P6) participants' recalls of episode event 22.  Sections of recall describing each each episode event in Panel E are highlighted with the corresponding color.}
  \label{fig:precision-detail}
\end{figure}

@@

\subsection*{Experimental design and data collection}
Data were collected by \cite{ChenEtal17}.  In brief, participants ($n=22$) viewed the first 48 minutes of ``A Study in Pink,'' the first episode of the BBC television show \textit{Sherlock}, while fMRI volumes were collected (TR = 1500~ms).  Participants were pre-screened to ensure they had never seen any episode of the show before.  The stimulus was divided into a 23~min (946~TR) and a 25~min (1030~TR) segment to mitigate technical issues related to the scanner.  After finishing the clip, participants were instructed to \citep[quoting from][]{ChenEtal17} ``describe what they recalled of the [episode] in as much detail as they could, to try to recount events in the original order they were viewed in, and to speak for at least 10 minutes if possible but that longer was better. They were told that completeness and detail were more important than temporal order, and that if at any point they realized they had missed something, to return to it. Participants were then allowed to speak for as long as they wished, and verbally indicated when they were finished (e.g., `I’m done').''  Five participants were dropped from the original dataset due to excessive head motion (2 participants), insufficient recall length (2 participants), or falling asleep during stimulus viewing (1 participant), resulting in a final sample size of $n=17$.  For additional details about the experimental procedure and scanning parameters, see \cite{ChenEtal17}.  The experimental protocol was approved by Princeton University's Institutional Review Board.

>>>>

\cite{ChenEtal17} collected hand-annotated information about each of 1,000 (manually identified) scenes spanning the roughly 50 minute video used in their study.  This information included: a brief narrative description of what was happening, the location where the scene took place, the names of any characters on the screen, and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the vocabulary for the topic model.  We then concatenated the sets of words across all features contained in overlapping sliding windows of (up to) 50 scenes, and treated each window as a single document for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the episode (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  We note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06}, in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched (interpolated) the resulting windows-by-topics matrix to match the time series of the 1,976 fMRI volumes collected as participants viewed the episode.

@@

\begin{figure}[tp]
  \centering
  \vspace*{-1cm}
  \includegraphics[width=.68\textwidth]{figs/precision_distinctiveness_detail}
  \caption{\small \textbf{Precision reflects the completeness of recall, whereas distinctiveness reflects recall specificity.} \textbf{A.} Recall precision by episode event.  Grey violin plots display kernel density estimates for the distribution of recall precision scores for a single episode event.  Colored dots within each violin plot represent individual participants' recall precision for the given event.  episode events are ordered along the $x$-axis by the average precision with which they were remembered.  \textbf{B.} The set of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising an example episode event (22) identified by the HMM.  Each action or feature is highlighted in a different color.  \textbf{C.} Excerpts from the most precise (P17) and least precise (P6) participants' recalls of episode event 22.  Descriptions of specific actions or features reflecting those highlighted in Panel B are highlighted in the corresponding color.  The text highlighted in gray denotes a (rare) false recall.  The unhighlighted text denotes correctly recalled information about other episode events.  \textbf{D.} Recall distinctiveness by episode event.  Kernel density estimates for each episode event's distribution of recall distinctiveness scores, analogous to Panel A.  \textbf{E.} The sets of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising episode events described by the example participants in Panel F.  Each event's text is highlighted in a different color.  \textbf{F.} The sentences comprising the most distinctive (P9) and least distinctive (P6) participants' recalls of episode event 22.  Sections of recall describing each each episode event in Panel E are highlighted with the corresponding color.}
  \label{fig:precision-detail}
\end{figure}

@@

\subsection*{Paradigm and data collection}
Data were collected by \cite{ChenEtal17}.  In brief, participants ($n=22$) viewed the first 48 minutes of ``A Study in Pink,'' the first episode of the BBC television show \textit{Sherlock}, while fMRI volumes were collected (TR = 1500~ms).  Participants were pre-screened to ensure they had never seen any episode of the show before.  The stimulus was divided into a 23~min (946~TR) and a 25~min (1030~TR) segment to mitigate technical issues related to the scanner.  After finishing the clip, participants were instructed to \citep[quoting from][]{ChenEtal17} ``describe what they recalled of the [episode] in as much detail as they could, to try to recount events in the original order they were viewed in, and to speak for at least 10 minutes if possible but that longer was better. They were told that completeness and detail were more important than temporal order, and that if at any point they realized they had missed something, to return to it. Participants were then allowed to speak for as long as they wished, and verbally indicated when they were finished (e.g., `I’m done').''  Five participants were dropped from the original dataset due to excessive head motion (2 participants), insufficient recall length (2 participants), or falling asleep during stimulus viewing (1 participant), resulting in a final sample size of $n=17$.  For additional details about the testing procedures and scanning parameters, see \cite{ChenEtal17}.  The testing protocol was approved by Princeton University's Institutional Review Board.
